\documentclass[11pt]{article}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{algorithmic}
\usepackage{relsize}
\usepackage{algorithm}
\usepackage{todonotes}
\usepackage{url}
\usepackage{tikz}
\usetikzlibrary{arrows}

\usepackage{listings}
 \lstset{
            language=Matlab,                                % choose the language of the code
    %       basicstyle=10pt,                                % the size of the fonts that are used for the code
            numbers=left,                                   % where to put the line-numbers
            numberstyle=\footnotesize,                      % the size of the fonts that are used for the line-numbers
            stepnumber=1,                                           % the step between two line-numbers. If it's 1 each line will be numbered
            numbersep=5pt,                                  % how far the line-numbers are from the code
    %       backgroundcolor=\color{white},          % choose the background color. You must add \usepackage{color}
            showspaces=false,                               % show spaces adding particular underscores
            showstringspaces=false,                         % underline spaces within strings
            showtabs=false,                                         % show tabs within strings adding particular underscores
    %       frame=single,                                           % adds a frame around the code
    %       tabsize=2,                                              % sets default tabsize to 2 spaces
    %       captionpos=b,                                           % sets the caption-position to bottom
            breaklines=true,                                        % sets automatic line breaking
            breakatwhitespace=false,                        % sets if automatic breaks should only happen at whitespace
            escapeinside={\%*}{*)}                          % if you want to add a comment within your code
}

\setlength{\paperwidth}{8.5in}
\setlength{\paperheight}{11in}
\setlength{\voffset}{-0.2in}
\setlength{\topmargin}{0in}
\setlength{\headheight}{0in}
\setlength{\headsep}{0in}
\setlength{\footskip}{30pt}
\setlength{\textheight}{9.25in}
\setlength{\hoffset}{0in}
\setlength{\oddsidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\parindent}{0in}
\setlength{\parskip}{9pt}

\newcommand{\ben}{\begin{enumerate}}
\newcommand{\een}{\end{enumerate}}

\DeclareGraphicsRule{.JPG}{eps}{*}{`jpeg2ps #1}

\title{Maximizing Locality of Graph Traversals}
\author{Casey Battaglino}
\date{}
\begin{document}
\maketitle
%\tableofcontents

\section{Breadth-First Traversal}

Given a graph $G$ with adjacency matrix $A \in \mathbb{R^{(N \times N)}}$, a frontier expansion of a breadth-first search is equivalent to the sparse matrix-vector product $A^Tx$, where $w$ contains a nonzero for each element in the current reach. 

If $A$ is represented in CSR format, the number of flops required is $2 \times nnz(A)$. However, only those rows in $A$ that correspond to nonzero indices in $x$ need be loaded from memory. 

\subsection{}
Locate BFS optimization models along a spectrum of run-time-conscious $\to$ theory-conscious. (Iteratively optimizing bottlenecks vs. attempting to minimize theoretical 

\subsection{Replacement Policies}
The LRU buffer replacement strategy is described as a stack in~\cite{Dan:1990:AAL:98457.98525}, where a requested item not found in the stack is moved to the top of the stack, and the least recently used item (the bottom of the stack), is removed. If an item is requested that is in the stack, it is removed from the stack and placed at the top. 

\subsection{Tuning}
\cite{Xia_topologicallyadaptive}

\subsection{Cache Model}
When a row is loaded from memory, its nonzeros are stored in cache according to some cache policy. Assume for now that we are computing on a single core with multiple levels of cache. After a particular traversal step $i$, we can imagine a permutation $p_i \in \Sigma_n$ that orders the rows of $A$ in the order in which they are stored in cache. 

Now consider the traversal step $i+1$: based on the particular cache policy, we can imagine a recurrence relation that describes $p_{i+1}=f(p_i; A, x_i)$

\subsection{Notes}
\begin{enumerate}
\item What does the spectrum tell us about the growth of $x_i$?
\item What does the degree distribution tell us (gen function methodology)?
\item Kronecker - PL mapping
\end{enumerate}




\bibliographystyle{plain}
\bibliography{bib}


\end{document}
